Paper Report

### 2017-11-14 Paper_report.tex
Alignment score, non-probability

\item Purpose: Find $\theta$ which maximize p(X $|\theta$)
\item How: Use Generalized EM to estimate $\theta$
	\begin{enumerate}
	\item : Maximizing $p(X)$ means
	\begin{equation}
	p(x) = \sum_{z} p(X,Z) = E_{q(Z)}[\ln p(X,Z|\theta)] + H[q(Z)] + KL(q(Z) || p(Z|X,\theta))
	\end{equation}
	\item E-step: Estimate $p(Z|X)$ based on current $\theta$
	\item M-step: Consider $q(Z) = p(Z|X)$ \\
	$KL(q(Z) || p(Z|X,\theta)) = 0$ \\
	$H[q(Z)]$ constant
	\item Maximizing $p(X)$ becomes maximizing $E_{q(Z)}[\ln p(X,Z|\theta)]$
	\end{enumerate}
\item Result: Gradient of the joint probability scaled by the posteriors
	\begin{equation}
	J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)
	\label{GradientJointProbability} 
	\end{equation}
\item Problem: How to calculate $p(X,Z)$ ?
\end{itemize}

@@

\item Therefore, the probability of a given sequence of observation X and latent variables Z (Factorization of the joint probability):
	\begin{equation}
	p(X,Z) = \prod_{t=1}^{n + 1} p(z_{t}|z_{t-1})\prod_{t=1}^{n} p(x_{t}|z_{t})
	\label{ProbabilityHMM}
	\end{equation}

\item Result: Combine $p(X,Z)$ \eqref{ProbabilityHMM} and gradient $J(\theta)$ \eqref{GradientJointProbability}

@@

\begin{equation}
s_{NN}(a|e, f) = \prod_{i = 1}^{|e|} t_{lex}(e_{i},f_{a_{i}}|e,f) t_{d}(a_{i},a_{i-1}|e,f)
\end{equation}
	\begin{itemize}
	\item $f_{a_{i}}$ represents not only this word but the context around this word. Surrounding words of both source and target word are input of DNN, which is handling the context of both sides. This could reduce the explosion of parameter number. \\ In this case, they use fixed length windows surrounding both $e_{i}$ and $f_{j}$. 
	\item $t_{lex}$: Lexical translation score.
	\begin{equation}
	t_{lex}(e_{i},f_{j}|e,f) = functions_{NN}(window(e_{i}),window(f_{j}) )
	\end{equation}
	\item $t_{d}$: Distortion score.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1}|window(f_{a_{i-1}})) = functions_{NN}(window(f_{a_{i-1}}))
	\end{equation}
	They recognize that this lexicalized distortion does not produce a better alignment. They reverse to the simple version.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1})
	\end{equation}
	\end{itemize}

\subsection{Result}

@@

\subsection{Main idea}
This research is mainly based on \citep{Yang13word}. The difference is that the alignment score is calculated in recurrent approach RNNs, which means that $a_j$ depends on all previous position $a_{0...j-1}$.

@@

@@

\section{Overview about statistical alignment \cite{Och2003Systematic}}
\subsection{General problem}
Input:
\begin{itemize}
\item A source sentence: $f_1^J = f_1,...,f_j,...f_J$
\item A target sentence: $e_1^I = f_1,...,f_i,...f_I$
\end{itemize}
Output: Alignment map $j \rightarrow i = a_j$
\begin{itemize}
\item An alignment: $a_1^J = a_1,...,a_j,...a_J$
\end{itemize}

\subsubsection{How to model a word alignment ?}
Our work is modelling the relationship between a source sentence and a target sentence. \\
We start from the view of translation model, which is finding this best translation $e_1^I$ for a source sentence $f_1^J$.
\begin{equation}
e_1^I = argmax_{e_1^I} p(e_1^I|f_1^J) = argmax_{e_1^I} \frac{p(f_1^J|e_1^I)p(e_1^I)}{p(f_1^J)}
\end{equation}

@@

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\label{AlignmentModelEquation}
\end{equation}

@@

\section{Word alignment model in Neural Network}
We start from this formula \ref{AlignmentModelEquation}:

@@

\subsection{Non-probability approach, we use score}
The papers include \cite{Yang13word}, \cite{Tamura14recurrent}.

>>>>

\item Purpose: Find $\theta$ which maximize p(X $|\theta$)
\item How: Use Generalized EM to estimate $\theta$
	\begin{enumerate}
	\item : Maximizing $p(X)$ means
	\begin{equation}
	p(x) = \sum_{z} p(X,Z) = E_{q(Z)}[\ln p(X,Z|\theta)] + H[q(Z)] + KL(q(Z) || p(Z|X,\theta))
	\end{equation}
	\item E-step: Estimate $p(Z|X)$ based on current $\theta$
	\item M-step: Consider $q(Z) = p(Z|X)$ \\
	$KL(q(Z) || p(Z|X,\theta)) = 0$ \\
	$H[q(Z)]$ constant
	\item Maximizing $p(X)$ becomes maximizing $E_{q(Z)}[\ln p(X,Z|\theta)]$
	\end{enumerate}
\item Result: Gradient of the joint probability scaled by the posteriors
	\begin{equation}
	J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)
	\label{Tran16unsupervisedGradientJointProbability} 
	\end{equation}
\item Problem: How to calculate $p(X,Z)$ ?
\end{itemize}

@@

\item Therefore, the probability of a given sequence of observation X and latent variables Z (Factorization of the joint probability):
	\begin{equation}
	p(X,Z) = \prod_{t=1}^{n + 1} p(z_{t}|z_{t-1})\prod_{t=1}^{n} p(x_{t}|z_{t})
	\label{Tran16unsupervisedProbabilityHMM}
	\end{equation}

\item Result: Combine $p(X,Z)$ \eqref{Tran16unsupervisedProbabilityHMM} and gradient $J(\theta)$ \eqref{Tran16unsupervisedGradientJointProbability}

@@

\begin{equation}
s_{NN}(a|e, f) = \prod_{i = 1}^{|e|} t_{lex}(e_{i},f_{a_{i}}|e,f) t_{d}(a_{i},a_{i-1}|e,f)
\label{Yang13wordNNAlignmentScore}
\end{equation}

	\begin{itemize}
	\item $s_{NN}(a|e, f)$ is the score of an alignment based on the source and target sentence.
	\item $f_{a_{i}}$ represents not only this word but the context around this word. Surrounding words of both source and target word are input of DNN, which is handling the context of both sides. This could reduce the explosion of parameter number. \\ In this case, they use fixed length windows surrounding both $e_{i}$ and $f_{j}$. 
	\item $t_{lex}$: Lexical translation score.
	\begin{equation}
	t_{lex}(e_{i},f_{j}|e,f) = functions_{NN}(window(e_{i}),window(f_{j}) )
	\label{Yang13wordLexicalTranslationScore}
	\end{equation}
	\item $t_{d}$: Distortion score.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1}|window(f_{a_{i-1}})) = functions_{NN}(window(f_{a_{i-1}}))
	\end{equation}
	They recognize that this lexicalized distortion does not produce a better alignment. They reverse to the simple version.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1})
	\end{equation}
	\end{itemize}

\subsection{Loss in supervised learning}
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - s_{\theta}(a^+ |\textbf{e},\textbf{f}) + s_{\theta}(a^- |\textbf{e},\textbf{f}) )
\label{Yang13wordAlignmentLoss}
\end{equation}
where
\begin{itemize}
\item $a^+$: Gold alignment path
\item $a^-$: Highest scoring incorrect alignment path under $\theta$
\item $s_\theta$: Score for alignment path defined in \eqref{Yang13wordNNAlignmentScore}
\end{itemize}

\subsection{Analysis}
\subsubsection{Two loss functions during training}
They pre-train word embedding with monolingual data.
They train firstly NN with the loss from lexical translation score and then the loss \eqref{Yang13wordAlignmentLoss} because they recognize that using only the loss \eqref{Yang13wordAlignmentLoss} is not efficient.

The loss of lexical translation score:
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - t_{lex, \theta}((e,f)^+ | \textbf{e},\textbf{f}) + t_{lex, \theta}((e,f)^- | \textbf{e},\textbf{f}) )
\end{equation}
where
\begin{itemize}
\item $(e,f)^+$: Correct word pair
\item $(e,f)^-$: Incorrect word pair
\item $t_{lex}$: Score of \eqref{Yang13wordLexicalTranslationScore}
\end{itemize}

\subsubsection{Result}
Baselines are HMM and IBM4. IBM4+NN takes the first place, followed by IBM4 and then HMM+NN.

\subsubsection{Notes}
They conclude also that the size of window influence on the accuracy of translation score.
The large number of hidden layer does not return any improvement.

@@

\subsection{Main idea}
This research is mainly based on \citep{Yang13word}. The difference is that the score of alignment is calculated in recurrent approach RNNs, which means that $a_j$ depends on all previous position $a_{0...j-1}$.
The score of an alignment in \citep{Yang13word} has two components: One is for lexical translation and the other is for alignment. This research proposes a single score.

@@

The equation of \citep{Yang13word} (Reverse the source and the target sentence)
\begin{equation}
s_{NN}(a_1^ J|e_1^ I, f_1^ J) = \prod_{j = 1}^{J} t_{d}(a_{j},a_{j-1}|window(e_{a_{j-1}})) *t_{lex}(f_{j},e_{a_{j}}|window(e_{a_{j}}), window(f_{j}))
\end{equation}

It is modified for RNNs:

\begin{equation}
s_{NN}(a_1^ J|e_1^ I, f_1^ J) = \prod_{j = 1}^{J} t_{RNN}(a_{j}|a_1^{j-1},e_{a_{j}},f_j  )
\end{equation}

\subsection{Loss in unsupervised learning} 
\begin{equation}
loss(\theta) = max(0 , 1 - \sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\label{Tamura14recurrentLossUnsupervisedLearning}
\end{equation}
where
\begin{itemize}
\item T: Training data (Bilingual sentences) as Observed data
\item $\Omega$ : Full translation search space as neighbourhood of observe data, .
\item $\Phi$ : Set of all possible alignments given (\textbf{e},\textbf{f})
\item $E_{\Phi} [s_{\theta}]$ : Expected value of the scores $s_\theta$ on $\Phi$
\item $e^+$ : A target language sentence in T
\item $e^-$ : A pseudo-target language sentence
\item $\sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]$ : Expectation term for observed data
\item$ \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )$ : Expectation term for neighbourhood
\end{itemize}

They would like to reduce the computation, they randomly select N pseudo-target language sentences for each $f^+$. \eqref{Tamura14recurrentLossUnsupervisedLearning} becomes:
\begin{equation}
loss(\theta) = \sum_{\textbf{f}^+ \in T} max(0 , 1 -  E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \frac{1}{N} \sum_{\textbf{e}^-} E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\end{equation}
where
\begin{itemize}
\item $\textbf{e}^-$ : a pseudo-target language sentence with the same length $|\textbf{e}^-| = |\textbf{e}^+|$
\item GEN: A subset of all possible alignment generated by beam search
\end{itemize}

\subsection{Agreement constraints} 
\subsection{Analysis}

@@

\section{Overview about statistical alignment }

\subsection{Word-based alignment ? \cite{Och2003Systematic}}
Input:
\begin{itemize}
\item A source sentence: $f_1^J = f_1,...,f_j,...f_J$
\item A target sentence: $e_1^I = f_1,...,f_i,...f_I$
\end{itemize}
Output: Alignment map $j \rightarrow i = a_j$
\begin{itemize}
\item An alignment: $a_1^J = a_1,...,a_j,...a_J$
\end{itemize}

Our work is modelling the relationship between a source sentence and a target sentence. \\
We start from the view of translation model, which is finding this best translation $e_1^I$ for a source sentence $f_1^J$.
\begin{equation}
e_1^I = argmax_{e_1^I} p(e_1^I|f_1^J) = argmax_{e_1^I} \frac{p(f_1^J|e_1^I)p(e_1^I)}{p(f_1^J)}
\end{equation}

@@

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\label{Och2003SystematicAlignmentModelEquation}
\end{equation}

@@

\section{Word alignment model in Neural Network}
We start from this formula \eqref{Och2003SystematicAlignmentModelEquation}:

@@

\subsection{Probability approach}
The papers include \cite{Tran16unsupervised}.

\subsection{Non-probability - Score approach}
The papers include \cite{Yang13word}, \cite{Tamura14recurrent}.
